Fast, Scalable Phrase-Based SMT Decoding

### 2016-02-04 acl.2016/acl2016.tex
acl paper

We also note that distortion limit which constrains hypothesis extension is dependent on the hypothesis' coverage vector, $C$ \emph{and} the end position of most recent source word that has been translated, $e$. The distortion limit must be checked for every instance of a hypothesis and translation rule, Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack}.
\begin{figure} [h]
\begin{algorithmic}
%\REQUIRE source sentence $s$, translation options
\FORALL{ $hypo$ in $stack_{|C|}$} 
  \FORALL{ translation rules}
      \IF {can-expand($C(hypo)$, $e(hypo)$, translation rule range)}
	\STATE expand hypo with translation rule $\rightarrow$ new hypo
	\STATE add new hypo to next stack
      \ENDIF
  \ENDFOR % { translation options}
\ENDFOR % { hypothesis in $stack_i$} 
\end{algorithmic}
\caption{Hypothesis Expansion with Unsorted Stack}
\label{algo:Hypothesis Expansion with Unsorted Stack}
\end{figure}
However, by separating hypotheses into \emph{'ministacks'} according to coverage and end position, the distortion limit only needs to be checked for each ministack, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}.
\begin{figure} [h]
\begin{algorithmic}
%\REQUIRE source sentence $s$, translation options
\FORALL{ $ministack_{C, e}$ in $stack_{|C|}$} 
  \FORALL{ translation rules}
    \IF {can-expand($C$, $e$, translation rule range)}
      \FORALL{ $hypo$ in $ministack_{C, e}$} 
	\STATE expand hypo with translation rule $\rightarrow$ new hypo
	\STATE add new hypo to next ministack
      \ENDFOR
    \ENDIF
  \ENDFOR % { translation options}
\ENDFOR % { hypothesis in $stack_i$} 
\end{algorithmic}
\caption{Hypothesis Expansion with Sorted Stack}
\label{algo:Hypothesis Expansion with Sorted Stack}
\end{figure}
We therefore looked at the effects of three stack configurations.
\begin{enumerate}
  \item coverage cardinality,
  \item coverage,
  \item coverage and end position of most recent translated source word.
\end{enumerate}
We therefore looked at the effects of three stack configurations.
\begin{enumerate}
  \item coverage cardinality,
  \item coverage,
  \item coverage and end position of most recent translated source word.
\end{enumerate}
Table~\ref{tab:stack-configuration} and Figure~\ref{fig:stack-configuration} present the decoding time and average model on the held out test set.
\begin{table}[h]
\tiny
\begin{center}
\begin{tabular}{|r|r|r|r|r|r|r|} \hline
Pop-limit	& \multicolumn{2}{c|}{Cardinality}	& \multicolumn{2}{c|}{Coverage}	& \multicolumn{2}{c|}{Coverage \& end pos} \\ \hline
		& Time		& Score			& Time		& Score		& Time		& Score	\\ \hline
100  		& 73		& -8.64513 		& 75		& -8.64513	& 72	& -8.64513 \\
500		& 237		& -8.59563		& 225		& -8.59563 	& 229	& -8.59612 \\
1,000	 	& 416		& -8.58700		& 397		& -8.58700	& 423	& -8.58700 \\ 
5,000 		& 1930		& -8.58165		& 1931		& -8.58165	& 2153	& -8.58098 \\ 
10,000 		& 3619		& -8.58133		& 3630		& -8.58133	& 4576	& -8.58015 \\ 
15,000 		& 4830		& -8.58130		& 5001		& -8.58130	& 7156	& -8.57999 \\ 
20,000 		& 5849		& -8.58130		& 5916		& -8.58130	& 9583	& -8.57994 \\ \hline
\end{tabular}
\end{center}
\caption{Decoding time (in secs with 32 threads) and average model scores for different stack configurations}
\label{tab:stack-configuration}
\end{table}

\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{stack-configuration.png}} 
\end{tabular}
\caption{Trade-off between decoding time average model scores for different stack configurations}
\label{fig:stack-configuration}
\end{figure}

The algorithms in Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack} and ~\ref{algo:Hypothesis Expansion with Sorted Stack} describes the standard phrase-based decoding algorithm, however, the stack configuration issues are the same in the cube-pruning algorithm. Furthermore, the cube-pruning algorithm also contain a priority queue which is attached to coverage cardinality in Moses, Figure~\ref{algo:Cube Pruning with Unsorted Stack}.
\begin{figure} [h]
\begin{algorithmic}
%\REQUIRE source sentence $s$, translation options
\STATE initialize $queue_{|C|}$
\FOR{1 to pop-limit}
  \STATE get best $item$ in $queue_{|C|}$
  \STATE create new hypo from $item$
  \STATE add new hypo to new $stack_{|C|}$
  \STATE create next $items$
  \STATE add new $items$ to $queue_{|C|}$
\ENDFOR % { hypothesis in $stack_i$} 
\end{algorithmic}
\caption{Cube Pruning with Unsorted Stack}
\label{algo:Cube Pruning with Unsorted Stack}
\end{figure}
We experiment with changing the queue, $queue_{|C|}$, by having separate queues for each (1) unique coverage $queue_{C}$, and each (2) unique coverage and end position $queue_{C, e}$. The stack configuration stayed constant.

>>>>

We also note that distortion limit which constrains hypothesis extension is dependent on the hypothesis' coverage vector, $C$ \emph{and} the end position of most recent source word that has been translated, $e$. The distortion limit must be checked for every instance of a hypothesis and translation rule, Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack}.
\begin{figure} [h]
\begin{algorithmic}
%\REQUIRE source sentence $s$, translation options
\FORALL{ $hypo$ in $stack_{|C|}$} 
  \FORALL{ translation rules}
      \IF {can-expand($C(hypo)$, $e(hypo)$, translation rule range)}
	\STATE expand hypo with translation rule $\rightarrow$ new hypo
	\STATE add new hypo to next stack
      \ENDIF
  \ENDFOR % { translation options}
\ENDFOR % { hypothesis in $stack_i$} 
\end{algorithmic}
\caption{Hypothesis Expansion with Cardinality Stacks}
\label{algo:Hypothesis Expansion with Unsorted Stack}
\end{figure}
However, by separating hypotheses into set of hypotheses (\emph{'ministacks'}) according to coverage and end position, the distortion limit only needs to be checked for each ministack, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}. Furthermore, stack pruning is done on each of these hypotheses set therefore, changing how hypotheses are grouped can affect model scores. 
\begin{figure} [h]
\begin{algorithmic}
%\REQUIRE source sentence $s$, translation options
\FORALL{ $ministack_{C, e}$ in $stack_{|C|}$} 
  \FORALL{ translation rules}
    \IF {can-expand($C$, $e$, translation rule range)}
      \FORALL{ $hypo$ in $ministack_{C, e}$} 
	\STATE expand hypo with translation rule $\rightarrow$ new hypo
	\STATE add new hypo to next ministack
      \ENDFOR
    \ENDIF
  \ENDFOR % { translation options}
\ENDFOR % { hypothesis in $stack_i$} 
\end{algorithmic}
\caption{Hypothesis Expansion with Coverage \& End Position Stacks}
\label{algo:Hypothesis Expansion with Sorted Stack}
\end{figure}
We therefore looked at the effects of three stack configurations.
\begin{enumerate}
  \item coverage cardinality,
  \item coverage,
  \item coverage and end position of most recent translated source word.
\end{enumerate}
We therefore looked at the effects of three stack configurations.
\begin{enumerate}
  \item coverage cardinality,
  \item coverage,
  \item coverage and end position of most recent translated source word.
\end{enumerate}
Table~\ref{tab:stack-configuration} and Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model at various pop-limits.
\begin{table}[h]
\tiny
\begin{center}
\begin{tabular}{|r|r|r|r|r|r|r|} \hline
Pop-limit	& \multicolumn{2}{c|}{Cardinality}	& \multicolumn{2}{c|}{Coverage}	& \multicolumn{2}{c|}{Coverage \& end pos} \\ \hline
		& Time		& Score			& Time		& Score		& Time		& Score	\\ \hline
100  		& 73		& -8.64513 		& 75		& -8.64513	& 72	& -8.64513 \\
500		& 237		& -8.59563		& 225		& -8.59563 	& 229	& -8.59612 \\
1,000	 	& 416		& -8.58700		& 397		& -8.58700	& 423	& -8.58700 \\ 
5,000 		& 1930		& -8.58165		& 1931		& -8.58165	& 2153	& -8.58098 \\ 
10,000 		& 3619		& -8.58133		& 3630		& -8.58133	& 4576	& -8.58015 \\ 
15,000 		& 4830		& -8.58130		& 5001		& -8.58130	& 7156	& -8.57999 \\ 
20,000 		& 5849		& -8.58130		& 5916		& -8.58130	& 9583	& -8.57994 \\ \hline
\end{tabular}
\end{center}
\caption{Decoding time (in secs with 32 threads) and average model scores for different stack configurations}
\label{tab:stack-configuration}
\end{table}

\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{stack-configuration.png}} 
\end{tabular}
\caption{Trade-off between decoding time average model scores for different stack configurations}
\label{fig:stack-configuration}
\end{figure} 
As can be seen, the model scores for all stack configurations are identical for low pop-limits parameters but grouping hypotheses into coverage \& end position produces higher model scores for higher pop-limits. It is also slower but the time/quality tradeoff is better overall with this stack configuration. For lower pop-limits this configuration is slightly slower, but not by much, therefore, we shall stick with this Configuration for the remainder of the paper.

The cube-pruning algorithm contain a further priority queue which is attached to hypotheses sets which can be independent of how hypotheses are grouped. In the Moses implementation, the priority queue is also attached to the coverage cardinality, Figure~\ref{algo:Cube Pruning with Unsorted Stack}.
\begin{figure} [h]
\begin{algorithmic}
%\REQUIRE source sentence $s$, translation options
\STATE initialize $queue_{|C|}$
\FOR{1 to pop-limit}
  \STATE get best $item$ in $queue_{|C|}$
  \STATE create new hypo from $item$
  \STATE add new hypo to new $stack_{|C|}$
  \STATE create next $items$
  \STATE add new $items$ to $queue_{|C|}$
\ENDFOR % { hypothesis in $stack_i$} 
\end{algorithmic}
\caption{Cube Pruning with Unsorted Stack}
\label{algo:Cube Pruning with Unsorted Stack}
\end{figure}
Again, we experiment with diffirent queue configurations, having separate queues for each cardinality, coverage, and coverage \& end position. The stack configuration remained constant (coverage \& end position with pop-limit of 400).
\begin{table}[h]
\small
\begin{center}
\begin{tabular}{|r|r|r|} \hline
Queue configuration		& Time		& Score \\ \hline
Cardinality			& 192		& -8.59922 \\
Coverage			&		& -8.58635 \\
Coverage \& end position	&		& -8.58263 \\ \hline
\end{tabular}
\end{center}
\caption{Decoding time (in secs with 32 threads) and average model scores for different queue configurations}
\label{tab:queue-configuration}
\end{table}
From the results in Table~\ref{tab:queue-configuration}, using finer grain queues results in better model scores, it is significantly slower to decode.

\subsection{Translation Model}
